---
title: "R_Introduction"
author: "Pei-Cheng LI (H24074085)"
date: "2020/3/12"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: readable
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Basic Operation
### Type
#### class: distinguish the type is character, numeric, integer, complex or logical(TRUE or FALSE)
```{r}
class("Hi")
class(2.5)
class(as.integer(2))
class(4+3i)
class(T)  # TRUE = T; FALSES = F
```
#### as.integer(x): let the x's type become integer
```{r}
as.integer(TRUE)        # logical 'TRUE' becomes integer
as.integer(FALSE)       # logical 'FALSE' becomes integer
class(as.integer("Hi")) # can not be a character
```

#### is.integer(x): is x integer?
```{r}
is.integer("Hi")
is.integer(as.integer(FALSE))
```

### Arithmetic 
#### Addition: +
```{r}
4 + 7
```
#### Subtraction: -
```{r}
4 - 7
```
#### Multiplication: *
```{r}
4 * 7
```
#### Division: /
```{r}
4 / 7
```

#### Power: ^ or **
```{r}
4 ^ 7    # 4 to the power of 7
4 ** 7   # 4 to the power of 7
```
#### Root: sqrt(x)
```{r}
sqrt(4)  # the square root of 4
```
#### Quotient: %/%
```{r}
4 %/% 7  # the quotient of 4 divided by 7
```
#### Remainder: %%
```{r}
4 %% 7   # the remainder of 4 divided by 7
```
#### sign(x): distinguish x is positive(recall 1), negative(recall -1) or zero(recall 0)
```{r}
sign(12)
sign(0)
sign(-12)
```

#### abs(x): take the absolute value of x
```{r}
abs(12)
abs(0)
abs(-12)
```

#### log(x): take the logarithm of x
```{r}
log(2)     # the logarithm with base e 
log1p(1)   # log(x) = log1p(x-1)
log(2, 10) # the logarithm with base 10
log10(2)   # the logarithm with base 10
```

#### exp(x): $e^x$
```{r}
exp(5)   
expm1(5) # expm1(x) = exp(x) - 1
```

#### ceiling(x): round up to an integer
```{r}
ceiling(2.43)
ceiling(3.5)
```

#### floor(x): round down to an integer
```{r}
floor(2.43)
floor(3.5)
```

#### trunc(x): extract the integer part
```{r}
trunc(2.43)
trunc(3.5)
```

#### round(x, digits = n): round x to the nearst nth
```{r}
round(123.456789)
round(2.43549812, digits = 3)
round(6457, digits = -2)
```

#### signif(x, digits = n): round x to the nearst (n-1)th
```{r}
signif(123.456789)
signif(2.43549812, digits = 3)
signif(6457, digits = 2)
```

#### Trigonometric Functions
```{r}
sin(0) # sine
cos(0) # cosine
tan(0) # tangent
```

#### Inverse Trigonometric Function
```{r}
asin(0) # arc-sine
acos(1) # arc-cosine
atan(0) # arc-tangent
```

#### Hyperbolic Trigonometric Function
```{r}
sinh(0) # hyperbolic sine 
cosh(0) # hyperbolic cosine
tanh(0) # hyperbolic tangent
```

#### Hyperbolic Inverse Trigonometric Function
```{r}
asinh(0) # hyperbolic arc-sine
acosh(1) # hyperbolic arc-cosine
atanh(0) # hyperbolic arc-tangent
```

## Variable and Data
### Variable
#### using <- to appoint the variable object
```{r}
A <- 5       # can be appoint as any types
a <- 8
A * a        # capital letter(A) and small letter(a) are different
A <- a <- 4  # A = a = 4
A - a
objects()    # show all objects in workspace (= ls())
rm(a)        # delete the object
exists("a")  # check whether the object exists 
```

#### NA(missing value), NULL(no meaning), Inf(infinite), NaN(not a number)
```{r}
is.na(c(0/0, NA))  # is NA?
is.nan(c(0/0, NA)) # is NaN?
length(NULL)       # the length of NULL is 0
Inf - Inf + 10     # the answer is 10?
NaN + Inf          # the answer is Inf?
NA + Inf           # the answer is Inf?
NaN + NA           # the answer is NaN?
```

### Vector
#### using c(...) to create a vector, the elements in same vector will become same type(character > complex > numeric > integer > logical)
```{r}
c('YA', TRUE, 5+6i)  # will become character
c(3+4i, FALSE, 2.7)  # will become complex
```

#### get the elements from vectors
```{r}
a <- c(red = 3, blue = 4, black = 5, yellow = 6)
a[1]        # get the factor's all information
a[[1]]      # get the factor's value
a['blue']   # get the factor's all information
a[['blue']] # get the factor's value
a[1:4]      # get many facotrs' information in one time 
```

#### x:y the number from x to y
```{r}
1:10
2:8
```

#### seq(s, e, by): the number from x to y 
```{r}
seq(1, 10)
seq(1, 10, 2)               # by = increment of sequence
seq(1, 10, length.out = 5)  # length.out = desired length of the sequence
```

#### rep(x, times, each): repeat x
```{r}
rep(8, 3)
rep(c('A','a'), 2)            # compare with next one
rep(c('A','a'), each = 2)     # the order is different
rep(c(2,4), c(3,5))
rep(seq(2, 10, 2), 2)         # compare with next one
rep(seq(2, 10, 2), rep(2, 5)) # the order is different
```

#### length(x): calculate how many elements in vector x
```{r}
length(c(1, 2, 3))
```

#### sum(x): the sum of all elements in vector x
```{r}
sum(c(1, 2, 3))
```

#### prod(x): the product of all elements in vector x
```{r}
prod(c(1, 2, 3))
```


#### cumsum(x): return the vector from the accumulated elements
```{r}
cumsum(c(1, 2, 3))
```

#### sort(x): return the sorted vector
```{r}
sort(c(3, 2, 1))
sort(c(1, 2 ,3), decreasing = TRUE)
```

#### order(x): return the sorted index vector
```{r}
order(c(6, 5, 4))
order(c(4, 6, 5), decreasing = TRUE)
```

#### rank(x):return the sorted index vector
```{r}
rank(c(6, 5, 4))
```


#### plus, subtraction, multiplication and division: the vectors need to be a multiple relationship
```{r}
c(2, 4) + c(1, 3, 5)
c(2, 4) - c(1, 3, 5, 7)
c(2, 4) * c(1, 3, 5)
c(2, 4) / c(1, 3, 5, 7)
```
### Array
#### array(x, c(a, b)): using x to make an array by a x b
```{r}
array(c(1, 0), c(2, 3))
```

#### rbind(x, y):  using rows merge into an array
```{r}
rbind(c(1, 2), c(3, 4))
```

#### cbind(x, y): using columns merge into an array
```{r}
cbind(c(1, 2), c(3, 4))
```

#### get the elements
```{r}
a <- rbind(c(5, 4, 3), c(2, 5, 0))
a
a[1,]     # get first row
a[,1]     # get first column
a[1, 1:2] # get the elements from first to second colums of  first row
```

#### plus, subtraction, multiplication and division: arrays need to have same size
```{r}
a + a
b <- rbind(c(1, 2), c(3, 4))
a - a*2
b * b # multiply corresponding array elements
b / b # divide corresponding array elements
```

#### length(x): count how many elements in an array
```{r}
length(a)
```

#### dim(x): the dimension of an array
```{r}
dim(a)
```

#### ncol(x) or nrow(x): count how many column or row in an array
```{r}
ncol(a)
nrow(a)
```

#### aperm(x): get the x's transporse
```{r}
aperm(a)
```

### Matrix
#### matrix(x, nrow = a, ncol = b): create a matrix
```{r}
matrix(c(1:4), nrow = 2, ncol = 2)               # fill in data according to column
matrix(c(1:4), nrow = 2, ncol = 2, byrow = TRUE) # fill in data according to row
```

#### get elements in the same way as array
```{r}
a <- rbind(c(5, 4, 3), c(2, 5, 0))
a
a[1,]     # get first row
a[,1]     # get first column
a[1, 1:2] # get the elements from first to second colums of  first row
```

#### t(x): transpose matrix
```{r}
t(a)
```

#### %*%: matrix multiplication
```{r}
a %*% t(a)
```

#### diag: generate a diagona motrix, or return the diagonal vector of the matrix
```{r}
b <- diag(c(rep(1, 3))) # generate a diagona matrix
b
diag(b)                 # return the diagonal vector of the matrix
```

#### det: calculate the determinant of a matrix, the matrix must be a symmetric matrix
```{r}
det(b)
```

#### solve: return the inverse of the matrix which is suitable for solving linear equations
```{r}
A <- rbind(c(1, 2), c(8, 7))
solve(A)    # solve x from Ax = b, A: A, b: Identity matrix
b <- rbind(c(1, 1), c(1, 1))
solve(A, b) # solve x from Ax = b, A: A, b: b
```

#### eigen: calculate the eigenvetor and eigenvalue
```{r}
eigen(A)
```

#### rownames: modify or check the row names
```{r}
rownames(A)
rownames(A) <- c('row1', 'row2')
rownames(A)
```

#### colnames: modify or check the column names
```{r}
colnames(A)
colnames(A) <- c('col1', 'col2')
colnames(A)
```

### Factor
#### like a vector after grading
```{r}
x <- c(5, 2, 0, 5, 2, 0)
factor(x)
factor(x, ordered = TRUE)                 # order factors
factor(x, labels = c("I", "Love", "You")) # name of level
factor(x, exclude = 5)                    # exclude something
```

#### get elements
```{r}
x[1] 
x[[1]] # same with using x[1]
x[1:3]
x[c(1, 4)]
```

#### as.factor: let variable become to factor
```{r}
as.factor(x)
```

#### is.factor: whether the variable is factor
```{r}
is.factor(x)
is.factor(as.factor(x))
```

#### is.ordered: whether the factor is ordered
```{r}
is.ordered(factor(x, ordered = TRUE))
is.ordered(factor(x, ordered = FALSE))
```

#### as.ordered: order factor
```{r}
as.ordered(factor(x))
```

#### which: find the eligible elements
```{r}
which(x == 5)
```

### List
#### be able to contain data with different data attributes
```{r}
x <- list(a = "Hi", b = FALSE, c = 46, d = c(8, 5))
x
```

#### get elements
```{r}
x$a
x[1]
x[[4]]
x[[4]][1]

```

#### as.list: create list
```{r}
as.list(c("Bye", "88"))
```

#### is.list: whether it is list
```{r}
is.list(x)
```

#### attributes/names: see all the names of elements
```{r}
attributes(x)
names(x)
```

### Dataframe
#### like a data sheet
```{r}
name <- c("J", "Q", "K")
height <- c(7, 9, 3)
weight <- c(4, 2, 8)
x <- data.frame(name, height, weight)
x
```

#### get elements
```{r}
x[2, ]
x[, 2]
x[2, 2]
x[, "name"]
x$name
```

#### head：get the top six data of the data frame (default is 6)
```{r}
head(x)     # because of not enough data, all are displayed
head(x, 1L) # only first column is displayed
```

#### names：modify or check the column names
```{r}
names(x)
names(x) <- c("n", "h", "w")
names(x)
```

#### colnames：set the column names
```{r}
colnames(x)
```

#### row.names：modify or check the row names
```{r}
row.names(x)
row.names(x) <- c("r1", "r2", "r3")
row.names(x) 
```

#### rownames：set the row names
```{r}
rownames(x)
```

#### summary：basic information of the data
```{r}
summary(x)
```

## Input and Output Data
### Input Data
```{r}
# read.table(file = "XXX.txt") 
# read.csv(file = "XXX.csv") 
```

### Output Data
```{r}
# write.table(data, file = "XXX.txt")
# write.csv(data, file = "XXX.csv")
```

### Change and Find Directory
```{r}
# setwd("directory") # change
# getwd()            # find
```

### Save and Load the Objets
```{r}
# source("XXX.R")                 # retrieve the objets in a future session
# save.image("XXX.RData")         # save all objets in the current workspace
# save(a,b,c, file = "XXX.RData") # save parts of objets
# oad("XXX.Rdata")                # retrieve data
```

### R Package
```{r}
# install.packages("XXX") # download a package
# library(XXX)            # call the package
```

## Process Control
### Logical Judgment
#### >, <, >=, <=, ==, != those are commen
```{r}
1 %in% c(1, 3)  # A %in% B: A in B?
x <- c(1, 3, 5) 
x > 0 && x == 0 # intersction: logical judgement to single value
x > 0 & x == 0  # intersction: logical judgement to vactor
x > 0 || x == 0 # union: logical judgement to single value
x > 0 | x == 0 # union: logical judgement to single value
```

### Conditional Execution
#### if, else if,  else
```{r}
x <- 2
if(x > 0){
  print("positive")
}else if(x < 0){
  print("negative")
}else{
  print("zero")
}
```

#### ifelse
```{r}
a <- ifelse(x/2 == 0, "even", "odd") # ifelse(statement, TRUE = "even", FALSE = "odd")
a
```

#### switch
```{r}
switch(3, "dog", "cat", "pig")                # switch(return value represents the number of program segments, program segment1, program segment2, ...)
switch("liebe", ich = 5, liebe = 2, dich = 0) # switch(return the name of the program segment , name1 = program segment, name2 = program segment, ...)
```

### Loop Structure
#### for
```{r}
for(x in c(1:3)){
  print(x^2)
}
```

#### while: if the statement is true, it will repeat
```{r}
x <- -1
while(x <= 0){
  print(x)
  x <- x+1  # it's important, or it'll become an infinite loop
}
```

#### repeat
#### break: out of the loop
#### next: skip this time's loop and do next time's
```{r}
x <- -5
repeat{
  if(x > 0){
    break
  }else if(x == 0){
    x <- x + 10
    next
  }
  x <- x + 1
  print(x)
}
x
```

## Define Function
### Define Function
#### using function(parameter){repeated execution of the program}
```{r}
times <- function(x, y){
  x^y
}
times(2, 10)
times(10, 2)
```

## Statistical Application
### Distribution
#### Binomial Distribution
$$ X \sim Bin(x,\ n,\ p)$$
$$ P(X=x) = (^{n}_{x})p^{x}(1-p)^{1-x},\ \ x=0, ..., n$$
```{r}
dbinom(10, 10, 0.5)   # dbinom(x, size, prob): the density corresponding to x
pbinom(4, 10, 0.5)    # pbinom(x, size, prob): cumulative probability corresponding to x
qbinom(0.52, 10, 0.5) # qbinom(p, size, prob): quantile corresponding to probability p
rbinom(10, 10, 0.5)   # rbinom(n, size, prob): generate n random numbers of binomial distribution
```

#### Poisson Distribution
$$ X \sim Poi(\lambda)$$
$$ P(X=x) = \frac{\lambda^{x}e^{-\lambda}}{x!},\ \ x=0, 1, 2, ...$$
```{r}
dpois(10, 10)  # dpois(x, lambda): the density corresponding to x
ppois(8, 10)   # ppois(x, lambda): cumulative probability corresponding to x
qpois(0.1, 10) # qpois(p, lambda): quantile corresponding to probability p
rpois(10, 10)  # rpois(n, lambda): generate n random numbers of poisson distribution
```

#### Continuous Uniform Distribution
$$ X \sim U(a,\ b)$$
$$ f(x) = \frac{1}{b-a}, \ \ a \leq x \leq b$$
```{r}
dunif(3, 0, 4)   # dunif(x, min, max): the density corresponding to x
punif(3, 0, 4)   # punif(x, min, max): cumulative probability corresponding to x
qunif(0.5, 0, 4) # qunif(p, min, max): quantile corresponding to probability p
runif(8, 0, 4)   # runif(n, min, max): generate n random numbers of uniform distribution
```

#### Gamma Distribution
$$ X \sim \Gamma(\alpha,\ \beta)$$
$$ f(x) = \frac{x^{\alpha-1}e^{{-x}/\beta}}{\Gamma(\alpha)\beta^{\alpha}}, \ \ x > 0$$
```{r}
dgamma(2, 5, 8)   # dgamma(x, shape, scale): the density corresponding to x
pgamma(2, 5, 8)   # pgamma(x, shape, scale): cumulative probability corresponding to x
qgamma(0.2, 5, 8) # qgamma(p, shape, scale): quantile corresponding to probability p
rgamma(2, 5, 8)   # rgamma(n, shape, scale): generate n random numbers of gamma distribution
```

#### Exponential Distribution
$$ X \sim Exp(\lambda)$$
$$ f(x) = \lambda e^{-x\lambda}, \ \ when \ x \geq 0$$
```{r}
dexp(6, 0.5)   # dexp(x, lambda): the density corresponding to x
pexp(6, 0.5)   # pexp(x, lambda): cumulative probability corresponding to x
qexp(0.6, 0.5) # qexp(p, lambda): quantile corresponding to probability p
rexp(6, 0.5)   # rexp(n, lambda): generate n random numbers of exponential distributjion
```

#### Chi-squared Distribution
$$ X \sim \chi^{2}(k)$$
$$ f(x) = \frac{x^{p/2-1}e^{{-x}/2}}{\Gamma(p/2)2^{p/2}}, \ \ x > 0$$
```{r}
dchisq(5, 4)   # dchisq(x, df): the density corresponding to x
pchisq(5, 4)   # pchisq(x, df): cumulative probability corresponding to x
qchisq(0.5, 4) # qchisq(p, df): quantile corresponding to probability p
rchisq(5, 4)   # rchisq(n, df): generate n random numbers of chi-square distribution
```

#### Normal Distribution
$$ X \sim N(\mu,\ \sigma^{2}) $$
$$ f(x) = \frac{1}{\sigma \sqrt{2\pi}}e^{-\frac{(x-\mu)^{2}}{2\sigma^{2}}}, \ \ -\infty < x < \infty  $$
```{r}
dnorm(0.05, 0, 1) # dnorm(x, mean, sd): the density corresponding to x
pnorm(0.05, 0, 1) # pnorm(x, mean, sd): cumulative probability corresponding to x
qnorm(0.05, 0, 1) # qnorm(x, mean, sd): quantile corresponding to probability p
rnorm(9, 0, 1)    # rnorm(x, mean, sd): generate n random numbers of normal distribution
```

#### Student t Distributioni
$$ X \sim t(r)$$
$$ f(x) = \frac{\Gamma[(r+1)/2]}{ \sqrt{\pi r}\Gamma(r/2)}\frac{1}{(1+x^2/r)^{(r+1)/2}}, \ \ -\infty < x < \infty $$
```{r}
dt(5, 2)   # dt(x, df): the density corresponding to x
pt(5, 2)   # pt(x, df): cumulative probability corresponding to x
qt(0.1, 2) # qt(p, df): quantile corresponding to probability p
rt(5, 2)   # rt(n, df): generate n random numbers of t distribution
```

#### F Distribution
$$ X \sim F(r_1,\ r_2)$$
$$ f(x) = \frac{\Gamma[(r_1+r_2)/2](r_1/r_2)^{r_1/2}}{\Gamma(r_1/2)\Gamma(r_2/2)}\frac{x^{r_1/2-1}}{(1+r_1x/r_2)^{(r_1+r_2)/2}}, \ \ -\infty < x < \infty  $$
```{r}
df(4, 2, 6)   # df(x, df1, df2): the density corresponding to x
pf(4, 2, 6)   # pf(x, df1, df2): cumulative probability corresponding to x
qf(0.4, 2, 6) # qf(p, df1, df2): quantile corresponding to probability p
rf(4, 2, 6)   # rf(n, df1, df2): generate n random numbers of F distribution
```

### Draw
#### histogram, boxplot, density plot
```{r}
x <- rnorm(100, 0, 1)
hist(x)
boxplot(x)
plot(density(x))
```

#### barchart, piechart
```{r}
x <- sample(c(1, 8), 10, replace = TRUE)
barplot(x) # for discrete variable
pie(x)     # for discrete variable
```

#### pairs
```{r}
a <- matrix(rnorm(30), 10, 3) # use matrix
pairs(a) 
```

#### plot
```{r}
x <- rnorm(100)
y <- 4*x + rnorm(100, 0, 1)
plot(x, y, main = "plot",      # main = title
     xlim = NULL, ylim = NULL, # range of an axis
     xlab = "x", ylab = "y",   # name of an axis
     col = "blue",             # col = color
     pch = 0                   # pch = symbol for points
     )
```

#### plot(..., type = ...)
```{r}
par(mfrow = c(2, 3)) # split plot ranges
plot(x, y, type = "p", main = "p")
plot(x, y, type = "l", main = "l")
plot(x, y, type = "b", main = "b")
plot(x, y, type = "o", main = "o")
plot(x, y, type = "h", main = "h")
plot(x, y, type = "n", main = "n")
```

#### plot(..., pch = ..., cex = ...)
```{r}
par(mfrow = c(2,3)) # split plot ranges 
plot(x, y, pch = 1, main = "pch = 1") 
plot(x, y, pch = 3, main = "pch = 3", cex.main = 2)   # cex.main = the size of the title
plot(x, y, pch = 10, main = "pch = 10", cex.axis = 2) # cex.axis = the size of the axis
plot(x, y, pch = 15, main = "pch = 15", cex = 2, cex.main = 2) 
plot(x, y, pch = 19, main = "pch = 19", cex = 2, cex.axis = 2) 
plot(x, y, pch = "#", main = "pch = #", cex = 2, cex.axis = 2, cex.lab = 2, cex.main = 2)
```

#### plot(..., type = “l”, lty =..., lwd = ...)
```{r}
t <- 1:20 
y <- rnorm(20)
par(mfrow = c(2,3)) # split plot ranges 
plot(t, y, type = "l") 
plot(t, y, type = "l", lty = 2)          # lty = the style of a line
plot(t, y, type = "l", lty = 2, lwd = 2) # lwd = the width of a line
plot(t, y, type = "l", lty = 3, lwd = 2) 
plot(t, y, type = "l", lty = 4, lwd = 2) 
plot(t, y, type = "l", lty = 5, lwd = 2)
```

#### plot(..., bty = ...)
```{r}
t <- 1:20 
y <- rnorm(20)
par(mfrow = c(2,3)) # split plot ranges 
plot(t, y, type = "l", bty = "o", main = "bty = \" o \"") # the box around the plot
plot(t, y, type = "l", bty = "n", main = "bty = \" n \"") 
plot(t, y, type = "l", bty = "7", main = "bty = \" 7 \"") 
plot(t, y, type = "l", bty = "L", main = "bty = \" L \"") 
plot(t, y, type = "l", bty = "C", main = "bty = \" C \"") 
plot(t, y, type = "l", bty = "U", main = "bty = \" U \"")
```

#### plot(..., col = ...)
```{r}
t <- 1:20 
y <- rnorm(20)
par(mfrow = c(2,3)) # split plot ranges 
plot(t, y, type = "l", col = 1) 
plot(t, y, type = "l", col = 2) 
plot(t, y, type = "l", col = "green") 
plot(t, y, type = "l", col = "grey") 
plot(t, y, type = "l", col = "#009999") 
plot(t, y, type = "l", col = "#0000FF")
```

#### curve(expression, from = ..., to = ...)
```{r}
par(mfrow = c(2, 3)) 
curve(exp(-x), 0, 10) 
curve(dnorm(x), -10, 10) 
curve(x^2, -10, 10) 
curve(sin, 0, 10) 
curve(3*exp(-3*x), 0, 6) 
curve(5*x+2, 0, 10)
```

#### add points or lines or text to an existing plot
```{r}
x <- rnorm(50) 
y <- 3*x + rnorm(50, 0, 1) 
y1 <- 5*x + rnorm(50, 0, 2)

plot(x, y, type = "p", pch = 19) 
abline(a = 0, b = 3)

points(x, y1, col = 4, pch = 20) 
abline(a = 0, b = 5, col = 4)

text(-1.5, 2, "A", cex = 3) 
text(1, -5, "B", cex = 3, col = 2) 
legend("topleft", c("Y", "Y1"), col = c(1,4), lty = c(1, 1), cex = 1)
```